---
title: "OPP workflow"
subtitle: "A step-by-step tutorial on visualizing seabird foraging trips and key at-sea areas."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OPP workflow}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "90%",
  fig.align = "center",
  dpi = 300,
  comment = "#>"
)

library(OPPtools)
```

This package contains analysis tools for processing seabird tracking data collected under Canada's Oceans Protection Plan. The workflow aims to identify at-sea trips for central place foraging seabirds, then use those trips to estimate key areas of use while the birds are at sea. Key areas are estimated using two approaches: a traditional kernel density estimation and a Brownian Bridge Movement Model (BBMM) based kernel density estimation.

The core of this package is built around the existing R package `track2KBA`, and further documentation on `track2KBA` methodology can be found in [Lascelles et al. 2016](https://onlinelibrary.wiley.com/doi/full/10.1111/ddi.12411).

# Installation

The `OPPtools` package can be downloaded and installed from Github if you have previously downloaded the `devtools` package. The second package you will need for this workflow is `track2KBA`, which is available for download from CRAN.

```{r installation, eval = FALSE}
# Install OPPtools
# install.packages("devtools")
devtools::install_github('popovs/OPPtools')

library(OPPtools)
```

# Data preparation

The workflow in this package is designed to work smoothly with data stored on Movebank. In order to download data, you will need to provide your Movebank credentials. `OPPtools` has an optional convenience function, `opp_movebank_key`, that allows you to securely store an encrypted copy of your Movebank credentials. If you set your credentials this way, you will not be prompted to enter your Movebank credentials further by any future function calls.

```{r set movebank credentials, eval = FALSE}
opp_movebank_key(username = "my_movebank_username") # You will then be prompted to enter your password
```

We can then download data directly off of Movebank. If you stored your credentials using `opp_movebank_key`, the data will be downloaded directly. Otherwise, you will be prompted to enter your Movebank username and password in the console.

```{r movebank download, eval = FALSE}
murres <- opp_download_data(248994009)
```

In these examples, data for *thick-billed murres* from Coats Island are also already bundled with the package.

```{r murres data}
data(murres) # same output as `opp_download_data(248994009)`

knitr::kable(head(murres)) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

# Explore raw tracks

The `opp_map` function allows you to quickly plot up your raw Movebank tracks to visualize them. The orange dot represents the Movebank project site, which in the case of our seabirds is the same as the colony location.

```{r plot raw tracks}
opp_map(murres)
```

Another useful thing to examine prior to continuing the workflow is a simple plot of time traveled vs. distance from colony for each bird. Looking at this will allow you to make informed decisions on minimum and maximum distances traveled during individual foraging trips in addition to time elapsed during foraging trips.

```{r plot time vs distance}
opp_explore_trips(murres)
```

# Define trips

From the above track exploration, we can extract reasonable trip parameters for the next stage of the workflow: the `opp_get_trips` function. It seems that most trips are characterized by a minimum distance of at least 2 km away from the colony. Additionally, a distance of 15 km seems reasonable to capture any trips where the bird went offshore but the logger than died (i.e., incomplete trips). This 15 km value will be used by `opp_get_trips` to mark any incomplete trips. Finally, based on these plots, it looks like we want to filter out any tracks that spend less than 2 hours away from the colony. In short, we're saying that any tracks that achieve of a distance of >2 km away from the colony for at least 2 hours can qualify as a 'trip'.

We also need to trim down the data using the `opp2KBA` function.

```{r trips params}
innerBuff = 2 # (km) minimum distance from the colony to be in a trip
returnBuff = 15 # (km) outer buffer to capture incomplete return trips
duration = 2 # (hrs) minimum trip duration

murres <- opp2KBA(murres)
```

## Extract trips

The `opp_get_trips` function will return a SpatialPointsDataFrame containing the GPS track locations for each bird, grouped by a trip ID. It will also label all GPS points deemed not to be part of a trip. If `showPlots = TRUE` (the default), plots visualizing the trips will be output.

```{r trips}
trips <- opp_get_trips(murres, 
                       innerBuff = innerBuff, 
                       returnBuff = returnBuff,
                       duration = duration)
```

# Interpolate tracks

One of the primary end-data products of this workflow is a kernel density estimate of seabird space use. However, a key statistical assumption of kernel densities is an even sampling regime. In the case of GPS points, this would mean an even gap of time between each data point. The current data do not satisfy this assumption. To account for this, we must first *interpolate* the data. The OPPTools library uses the crawl interpolation method to do so. See `?ctcrw_interpolation` for more information.

The output plots for `ctcrw_interpolation` are similar to the ones above, except for the addition of green circles. The green circles indicate interpolated tracks. The function only interpolated tracks identified as "Complete" by `opp_get_trips` above. This can be adjusted with the `type` argument in the function. 

```{r crawl interpolate}
interp <- ctcrw_interpolation(trips,
                              site = murres$site,
                              type = "Complete",
                              timestep = "2 min")
```

# Summmarizing our trips

`OPPtools` comes with a function to quickly explore trip details. It accepts the output from either `opp_get_trips` or `ctcrw_interpolation`. If you run `sum_trips` on an interpolated output, the table will provide both the raw and interpolated number of GPS locations.

```{r trip summary}
# opp_get_trips output
knitr::kable(sum_trips(trips)) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# ctcrw_interpolation output
knitr::kable(sum_trips(interp)) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

# Calculate kernels

There are two mapping methods for exploring key marine sites for seabirds using OPPtools: the first is a traditional kernel density estimate, and the second is Brownian Bridge Movement Model (BBMM) kernel density estimate. The first output is summarized and presented with the `track2KBA` mapping approach, while the second is a simple BBMM 50% or 95% UD contour.

## 1. Traditional KDE

The traditional KDE workflow is adapted from the `track2KBA` workflow, which can be found in greater detail [here](https://cran.r-project.org/web/packages/track2KBA/vignettes/track2kba_workflow.html). 

The `track2KBA` workflow has been simplified and condensed within the `OPPtools` package. Kernel smoothers are calculated within the same function that generates the kernel itself. By default, the `href` method of calculating kernel smoothers is used. See `?opp_kernel` for further details.

```{r standard kernel}
kde <- opp_kernel(interp,
                  smoother = "href",
                  res = 1 # grid resolution in sq km
                  )
```

## 2. BBMM KDE

```{r bbmm kernel}
bb_kde <- opp_bbmm(interp,
                   res = 1)
```

# Delineating OPP sites

For the context of emergency response, OPP is interested in areas of high use (and therefore at high risk in the case of an oil spill). An initial assessment of such areas comes from the `track2KBA` workflow, which first estimates how *representative* a series of tracks is (i.e., how much a few individuals' movements represent the movements of the population as a whole). Using this representativeness estimate, derived by the `track2KBA::repAssess` function, the `track2KBA::findSite` function then estimates how many individuals of the source population use the surrounding area. For more information on the `track2KBA::repAssess` and `track2KBA::findSite` functions, see the `track2KBA` documentation.

```{r kba repr}
repr <- track2KBA::repAssess(tracks = interp$interp,
                             KDE = kde, 
                             iteration = 50,
                             levelUD = 95,
                             avgMethod = "weighted")
```

Representativeness derived by the `track2KBA::repAssess` function can then be used in combination with `track2KBA::findSite` to estimate space-use by the population at large. `OPPtools` comes with functions to map the results of `track2KBA::findSite`.

```{r track2kba findsites}
t2k_sites <- track2KBA::findSite(KDE = kde,
                                 represent = repr$out,
                                 levelUD = 95,
                                 popSize = 60000, 
                                 polyOut = TRUE)


```

For sites specific to the OPP context, the function `opp_sites` delineates areas of high use around the colony. Specifically, it pulls out population use contours from the stacked kernel density estimates. By default, `opp_sites` calculates areas where at least 10%, 50%, and 75% of the population are expected to occur.

```{r opp sites}
sites <- opp_sites(kernels = kde,
                 population = 60000,
                 repr = repr$out)

opp_map(sites)
```

# Miscellanea

## More fun with kernels

With either of our kernels, we can use the `ud_stack` function if we are interested in taking a look at a population-level raster. `ud_stack` accepts a `weights` argument if you want to weigh certain tracks as more important than others in your stack. In this example, we will weight our bbmm kernels by number of locations in a track. We want longer tracks (that go farther from the colony) to have more importance in our population level kernel.

```{r ud_stack}
# Using a combination of aggregate and sum_trips, 
# let's pull out the number of locations per animal
n_locs <- aggregate(interp_n_locs ~ ID, data = sum_trips(interp), sum)[[2]]

# Stack our bbmm kernels, weighted by n_locs
pop_bbmm <- ud_stack(bb_kde, n_locs)

# Might be nice to compare to non-weighted

# Now we can visualize
mapview::mapview(pop_bbmm)
```

A common workflow for rasterized UDs typically involves extracting the 50% and 99% volume contours. With the `ud_vol` function, we can specify which contours we want to extract and reclassify the raster to show our two contour values:

```{r ud vol}
# By default, the two volume levels are 50% and 99%, but these can be changed if desired.
pop_vols <- ud_vol(pop_bbmm)

raster::plot(pop_vols)
```


