

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(OPPtools)
library(dplyr)
library(flextable)

mb_login <- move::movebankLogin('allison.patterson', 'djYgXmbct87mYL4')



title <- dat$report_title
report_date <- as.character(Sys.Date())
project <- dat$mb_project_num
spp <- dat$spp
minDist <- dat$minDist
maxDist <- dat$maxDist
minDur <- dat$minDur
gapTime <- dat$gapTime
gapDist <-  dat$gapDist
interpolateGaps <- dat$interpolateGaps
timestep <- dat$timestep

study_data <- move::getMovebankStudy(study = project, login = mb_login)
dep_data <- move::getMovebankReferenceTable(study = project, login = mb_login)

my_data <- opp_download_data(study = c(project),
                             login = mb_login,
                             start_month = NULL,
                             end_month = NULL,
                             season = NULL
)

```
---
title: `r title`
author: "Allison Patterson and Sarah Popov"
date: `r report_date`
output: pdf_document
---

### Project summary

`r spp`s (*`r study_data$taxon_ids`*) were tracked at the breeding colony on 
Triangle Island, BC (`r study_data$main_location_lat` N,	`r abs(study_data$main_location_lat)` W). 
There have been `r study_data$number_of_deployments` GPS deployments on `r study_data$number_of_individuals` unique individuals between `r as.Date(study_data$timestamp_first_deployed_location)` and `r as.Date(study_data$timestamp_last_deployed_location)`.


Data are stored on Movebank in project `r as.character(project)`, with the study name:
`r study_data$name`.


```{r get_data, echo=FALSE, message=FALSE, warning=FALSE}

my_data <- opp_download_data(study = c(project),
                             login = mb_login,
                             start_month = NULL,
                             end_month = NULL,
                             season = NULL
)

my_track2kba <- opp2KBA(data = my_data)

```

#### Defining foraging trips

Foraging trips were identified as any continuous track at least `r minDist` km from the colony, which lasted for a minimum of `r minDur` hours. Trips were considered incomplete if the trips started or ended more than `r maxDist` km from the colony. Trips were considered 'gappy' if there were gaps between successive locations greater than `r gapTime` hours, and the bird travelled farther than `r gapDist` km during that time. 

The figures below show trips from each deployment, as defined by the criteria described above.

```{r trips, echo=FALSE, message=FALSE, warning=FALSE}
my_trips <- opp_get_trips(data = my_track2kba, 
                          innerBuff = minDist, 
                          returnBuff = maxDist,
                          duration  = minDur, 
                          gapLimit = 100, 
                          #missingLocs = gapPer,
                          gapTime = 2,
                          gapDist = 5,
                          showPlots = T, 
                          plotsPerPage = 4)
```

#### Trip summary

`r spp` foraging trips by trip type.

```{r summary, echo=FALSE, message=FALSE, warning=FALSE}

ft <- my_trips@data %>%
  filter(Type != 'Non-trip') %>% 
  group_by(ID, tripID, Type) %>% 
  mutate(
    duration = as.numeric(difftime(max(DateTime), min(DateTime), units = 'hours'))
  ) %>% 
  group_by(Type) %>%
  summarize(
    `Number of trips` = length(unique(tripID)),
    `Number of locations` = n(),
    `Median duration (hrs)` = round(median(duration), 1),
    `Median trip distance (km)` = round(median(max(ColDist))/1000, 1)
    ) %>% flextable::flextable(cwidth =1.25) 

ft
  
```


#### Track interpolation

Locations within complete, incomplete, and gappy trips were interpolated at a `r timestep` interval using a continuous time correlated random walk model. Figures below show the raw GPS locations (violet) overlain with the interpolated locations (green). Gaps in tracks during a trip, as defined above, were `r ifelse(interpolateGaps == F, 'not','')` interpolated.


```{r interpolation, echo=FALSE, message=FALSE, warning=FALSE}

my_interp <- ctcrw_interpolation(data = my_trips, 
                                site = my_track2kba$site,
                                type = c('Complete','Incomplete','Gappy'),
                                timestep = '10 min',
                                interpolateGaps = F,
                                showPlots = T,
                                theta = c(8,2),
                                quiet = TRUE
)

```


#### Trip summaries
This table provides summary statistics for each trip included in the analysis. Including a comparison of the number of GPS fixes relative to the interpolated locations.

```{r trip_summary, echo=FALSE, message=FALSE, warning=FALSE}

my_trip_summary <- OPPtools::sum_trips(my_interp) %>% 
  dplyr::mutate(duration = signif(duration, 3),
         max_dist_km = signif(max_dist_km, 3),
         departure = as.Date(departure)) %>% 
  dplyr::select(tripID, raw_n_locs, interp_n_locs, departure, duration, max_dist_km, complete) %>% 
  dplyr::rename(`Trip ID` = tripID,
         `GPS fixes` = raw_n_locs,
         `Interpolated locations` = interp_n_locs,
         `Start date` = departure,
         `Maximum distance (km)` = max_dist_km,
         `Duration (hrs)` = duration,
         Type = complete) %>% 
  flextable::flextable(cwidth =c(1.25, 1, 1, 1.5, 1, 1, 1)) 

n_tracks <- length(unique(my_interp$interp$ID))

```


```{r short report, echo=FALSE, message=FALSE, warning=FALSE}

# Report will end here if there are fewer than 20 tracks in the dataset
if (n_tracks < 20) knitr::knit_exit()
```

#### Calculating kernel density estimates

```{r kernel_ud, echo=FALSE, message=FALSE, warning=FALSE, fig.show='hide'}

h <- opp_href(my_interp$interp)

my_ud <- opp_kernel(data = my_interp,
                    interpolated = TRUE,
                    smoother = "href",
                    extendGrid = 30,
                    res = 1)


if (n_tracks > 10) {
  ra50 <- invisible((track2KBA::repAssess(tracks = my_interp$interp, 
                                          KDE = my_ud, 
                                          level = 50, 
                                          iteration = 2, 
                                          avgMethod = "weighted")))
  
  ra95 <- invisible((track2KBA::repAssess(tracks = my_interp$interp, 
                                          KDE = my_ud, 
                                          level = 95, 
                                          iteration = 2, 
                                          avgMethod = "weighted")))
}

```

Kernel density estimates were calculated for each individual track using the median *'href'* smoothing parameter, `r h` km. Track representativeness was calculated using the track2KBA package. For core use areas, with utilization distribution of 50%, tracks were `r signif(ra50$out, 3)`% representative. At the home range level, with utilization distribution of 95%, tracks were `r signif(ra95$out, 3)`% representative.

