

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(OPPtools)
library(dplyr)
library(flextable)

mb_login <- move::movebankLogin('allison.patterson', 'djYgXmbct87mYL4')



title <- dat$report_title
report_date <- as.character(Sys.Date())
project <- dat$mb_project_num
spp <- dat$spp
minDist <- dat$minDist
maxDist <- dat$maxDist
minDur <- dat$minDur
gapTime <- dat$gapTime
gapDist <-  dat$gapDist
interpolateGaps <- dat$interpolateGaps
timestep <- dat$timestep

study_data <- move::getMovebankStudy(study = project, login = mb_login)
dep_data <- move::getMovebankReferenceTable(study = project, login = mb_login)

my_data <- opp_download_data(study = c(project),
                             login = mb_login,
                             start_month = NULL,
                             end_month = NULL,
                             season = NULL
)

```
---
title: `r title`
author: "Allison Patterson and Sarah Popov"
date: `r report_date`
output: pdf_document
---

### Project summary

`r spp`s (*`r study_data$taxon_ids`*) were tracked at the breeding colony on 
Triangle Island, BC (`r study_data$main_location_lat` N,	`r abs(study_data$main_location_lat)` W). 
There have been `r study_data$number_of_deployments` GPS deployments on `r study_data$number_of_individuals` unique individuals between `r as.Date(study_data$timestamp_first_deployed_location)` and `r as.Date(study_data$timestamp_last_deployed_location)`.


Data are stored on Movebank in project `r as.character(project)`, with the study name:
`r study_data$name`.


```{r get_data, echo=FALSE, message=FALSE, warning=FALSE}

my_data <- opp_download_data(study = c(project),
                             login = mb_login,
                             start_month = NULL,
                             end_month = NULL,
                             season = NULL
)

my_track2kba <- opp2KBA(data = my_data)

```

#### Defining foraging trips

Foraging trips were identified as any continuous track at least `r minDist` km from the colony, which lasted for a minimum of `r minDur` hours. Trips were considered incomplete if the trips started or ended more than `r maxDist` km from the colony. Trips were considered 'gappy' if there were gaps between successive locations greater than `r gapTime` hours, and the bird travelled farther than `r gapDist` km during that time. 

The figures below show trips from each deployment, as defined by the criteria described above.

```{r trips, echo=FALSE, message=FALSE, warning=FALSE}
my_trips <- opp_get_trips(data = my_track2kba, 
                          innerBuff = minDist, 
                          returnBuff = maxDist,
                          duration  = minDur, 
                          gapLimit = 100, 
                          #missingLocs = gapPer,
                          gapTime = 2,
                          gapDist = 5,
                          showPlots = T, 
                          plotsPerPage = 4)
```

#### Trip summary

`r spp` foraging trips by trip type.

```{r summary, echo=FALSE, message=FALSE, warning=FALSE}

ft <- my_trips@data %>%
  filter(Type != 'Non-trip') %>% 
  group_by(ID, tripID, Type) %>% 
  mutate(
    duration = as.numeric(difftime(max(DateTime), min(DateTime), units = 'hours'))
  ) %>% 
  group_by(Type) %>%
  summarize(
    `Number of trips` = length(unique(tripID)),
    `Number of locations` = n(),
    `Median duration (hrs)` = round(median(duration), 1),
    `Median trip distance (km)` = round(median(max(ColDist))/1000, 1)
    ) %>% flextable::flextable(cwidth =1.25) 

ft
  
```


#### Track interpolation

Locations within complete, incomplete, and gappy trips were interpolated at a `r timestep` interval using a continuous time correlated random walk model. Figures below show the raw GPS locations (violet) overlain with the interpolated locations (green). Gaps in tracks during a trip, as defined above, were `r ifelse(interpolateGaps == F, 'not','')` interpolated.


```{r interpolation, echo=FALSE, message=FALSE, warning=FALSE}

my_interp <- ctcrw_interpolation(data = my_trips, 
                                site = my_track2kba$site,
                                type = c('Complete','Incomplete','Gappy'),
                                timestep = '10 min',
                                interpolateGaps = F,
                                showPlots = T,
                                theta = c(8,2),
                                quiet = TRUE
)

```
